### 2014-09-06 guarantee.tex
-xz

\begin{proof}
The proof consists of two parts: (1) the trace is feasible, and (2) the trace can be generated by the program. We sketch the proof as follows.
\begin{itemize}
\item  $\tau'$ is feasible. Researchers~\cite{chao,maximal} point out a trace is feasible iff it satisfies the sequential consistency, i.e., $\forall object\ o, \tau'\downarrow_{o}$ satisfies the serial specification of the object $o$. If  all the events considered are of the types in Table~\ref{Ta:syntax}, then the sequential consistency precisely means the following: (1) read-write consistency, i.e., each read event of a variable should contain the value written by the most recent write event, and (2) the synchronization consistency, e.g., the lock acquire and release of a lock should not be interleaved by the lock operations of the same lock, the begin event of a thread should follow the fork event of the parent thread. Our inter-thread data flow constraints guarantee the read-write consistency among shared locations. The SSA form encodes the read-write consistency among local variables. The synchronization consistency is also captured as constraints in our solver. 
\item $\tau$ can be generated by the program, i.e., for each thread $t$, $\tau'\downarrow_{t}$ can be generated by the corresponding thread code.
This claim requires that,  for any two adjacent events $e$ and $e'$ in  $\tau'\downarrow_{t}$, (1) if $e$ is not a branch, $inst^e$ and $inst^{e'}$ should be adjacent in the code; (2) if $e$ is a branch, $inst^e$ and $inst^{e'}$ should be adjacent and the $inst^e$ should be evaluated as the boolean value indicated by $inst^{e'}$. 
We derive the trace  through either concrete execution or symbolic execution, which satisfies the requirements and guarantees the correctness of the claim.
\end{itemize}
\end{proof}

@@

First, to the best of our knowledge, none of the existing techniques automatically explore the un-executed branches by relaxing the schedules, while ours enables this.
Second, assume we disable the exploration of un-executed branches and all techniques start with the same trace. Our technique still explores more traces. 
Figure~\ref{fig:discuss} illustrates the difference pictorially, where two vertical lines represent the progress of two threads and the circles with numbers denote the events with ids. 
Existing techniques require the reads to read the same values, therefore, enforce the schedule order $e_5 \rightarrow e_6$ and $e_2\rightarrow e_3$. We allow the reads to read different values from different writes as long as the feasibility is preserved. Comparatively, we  enforce a weaker schedule order  $e_4 \rightarrow e_6$ and $e_2\rightarrow e_3$, which allows more scheduling, e.g., the concurrent execution of $e_7$ (or some event after $e_7$) and $e_5$.

However, our technique does not identify the maximal set of traces from a single trace. The underlying reason is that, we still enforce the schedule order for preserving the base object, e.g., $e_2\rightarrow e_3$. It is possible to further relax the schedule so that $e_3$ reads from $e_1$, but this requires challenging reasoning of the field accesses as their base objects become unknown. We plan to explore this idea in future work.

>>>>

\begin{proof}
The proof consists of two parts: (1) the trace is feasible, and (2) the trace can be generated by the program. We sketch the proof as follows.
\begin{itemize}
\item  $\tau'$ is feasible. Researchers~\cite{chao,maximal} point out a 
trace is feasible iff it satisfies the sequential consistency, 
i.e., $\forall object\ o, \tau'\downarrow_{o}$ satisfies the serial 
specification of the object $o$. If  all the events considered are of 
the types in Table~\ref{Ta:syntax}, then the sequential consistency 
precisely means the following: (1) read-write consistency, i.e., each 
read event of a variable should contain the value written by the most 
recent write event, and (2) the synchronization consistency, e.g., the 
acquisition and release of a lock should not interleave with other 
lock operations of the same lock, the start event of a thread should 
follow the fork event of the parent thread. Our inter-thread data 
flow constraints guarantee the read-write consistency among shared 
locations. The SSA form encodes the read-write consistency among local 
variables. The synchronization consistency is also captured as constraints 
in our solver. 
\item $\tau$ can be generated by the program, i.e., for each 
thread $t$, $\tau'\downarrow_{t}$ can be generated by the corresponding 
thread code. This claim requires that,  for any two adjacent 
events $e$ and $e'$ in  $\tau'\downarrow_{t}$, (1) if $e$ is not a 
branch, $inst^e$ and $inst^{e'}$ should be adjacent in the code; (2) 
if $e$ is a branch, $inst^e$ and $inst^{e'}$ should be adjacent in terms
of control flow and the $inst^e$ should be evaluated as the boolean 
value implied by $inst^{e'}$. We derive the trace  through either 
concrete execution or precise modeling of unexplored branches, 
which satisfies the requirements and guarantees the correctness of the claim.
\end{itemize}
\end{proof}

@@

First, to the best of our knowledge, none of the existing techniques 
automatically explore the un-executed branches by relaxing the schedules, whereas \tool enables this.
Second, assume we disable the exploration of un-executed branches and all 
techniques start with the same trace. Our technique still explores more 
traces. Figure~\ref{fig:discuss} illustrates the difference, where two 
vertical lines represent the progress of two threads and the circles 
with numbers denote events with ids. 
Existing techniques require the reads to read the same values, and hence
enforce the schedule order $e_5 \rightarrow e_6$ and $e_2\rightarrow e_3$. 
We allow the reads to read different values from different writes as long as 
the feasibility is preserved. Comparatively, we enforce a weaker 
schedule order  $e_4 \rightarrow e_6$ and $e_2\rightarrow e_3$, which 
allows more scheduling, e.g., the concurrent execution of $e_7$ (or 
some event after $e_7$) and $e_5$.

However, \tool has not reached the maximum bound of sound exploration.
% does not identify the maximal set of traces 
%from a single trace. 
The underlying reason is that, we still enforce the schedule for 
preserving the base object, e.g., $e_2\rightarrow e_3$. It is possible 
to further relax the schedule so that $e_3$ reads from $e_1$, but 
this requires challenging reasoning of the field accesses as their base 
objects become uncertain. We plan to explore this idea in future work.
